---
title: "Data Prep"
output: html_notebook
---

```{r include=FALSE} 
# Other Modules we may want later ggplot2, randtests, eventInterval
library( tibble )
library( dplyr )
library( rio )
library( bit64 )
library( psych )
```

# Import Data from File

## Import message data from csv source.
```{r}
.msgw_in_import <- rio::import( "./Data/eqfa-msgw-in-0709.txt", format="csv" )
.msgw_in_import
```

# Prepare input data

- Remove "PartyDetailsDefinitionRequest" Messages (they all have duplicate arrival times _and_ `null` processing time)
- Normalize event `arrival_t` at $t_0 = 0$ Method: $t_i = t_i - t_0 \mbox{ for } i=0,...,n$.
- Spread out messages with duplicate arrival times; Method: $t_i = t_i + k \mbox{ for } i = 0,...,n \mbox{ for } k=0..d$ where $d = $ number of duplicates at $t_i$ 
- Calculate inter-message `delta_t` in nanoseconds
- Convert latency/processing time `PNQM Latency (ms)` to `duration` in nanoseconds
- Use `Msg Type` as new factor `type`
- Calculate message exit time `departure_t` in nanoseconds

```{r}
input_limit <- 1.2e11 # Two minutes in ns

event_data <- .msgw_in_import %>%
  filter( !`Msg Type` %in% c( "PartyDetailsDefinitionRequest" ) ) %>%
  { if( input_limit ) filter(  ., ts_ns < first(ts_ns) + input_limit ) } %>%
  mutate(
    t = ts_ns - first(ts_ns),
    duration = as.numeric( na_if( `PNQM Latency (ms)`, "null" ) ) * 1e6,
    type = as.factor( `Msg Type` )
  ) %>%
  group_by( t ) %>%
  mutate(
    dup_count = n() - 1,
    dup_number = row_number() - 1
  ) %>%
  select( t, duration, type, dup_count, dup_number ) %>%
  ungroup() %>%
  arrange( t ) %>%
  mutate( arrival_t = t + dup_number, delta_t = arrival_t - lag(arrival_t), departure_t = arrival_t + duration ) %>%
  select( arrival_t, delta_t, type, duration, departure_t )

event_data
```
Input data arrival time is Monotonic: `r all(.msgw_in_import$ts_ns == cummax(.msgw_in_import$ts_ns) )`.
Successfully loaded `r nrow( event_data )` observations from the first $input_limit$ nanoseconds of our event data.

## Vizualizations against Event Time 
```{r}
hist( as.numeric( event_data$arrival_t ), breaks=120, col="blue", xlab = "Event Time in ns over the first two minutes of data.", main="Histogram of Event Freq per Sec" )

par(mfrow=c( 4, 1 ) )#, mai=c(0.1, 0.1, 0.1, 0.1) )
stripchart( as.numeric( head( event_data$arrival_t, 10000 ) ), pch=19, cex=0.5, col="blue", xlab="Event Time in ns", main="1-D Stripchart of Event Time; first 10000 observations." )
stripchart( as.numeric( head( event_data$arrival_t, 1000 ) ), pch=19, cex=0.5, col="blue", xlab="Event Time in ns", main="1-D Stripchart of Event Time; first 1000 observations." )
stripchart( as.numeric( head( event_data$arrival_t, 100 ) ), pch=19, cex=0.5, col="blue", xlab="Event Time in ns", main="1-D Stripchart of Event Time; first 100 observations." )
stripchart( as.numeric( head( event_data$arrival_t, 10 ) ), pch=19, cex=0.5, col="blue", xlab="Event Time in ns", main="1-D Stripchart of Event Time; first 10 observations." )
par( mfrow=c(1, 1) )

plot( event_data$arrival_t, event_data$delta_t, col="blue", pch=19, xlab="Event Time in ns", ylab="Inter-Event Delta_t in ns", main="Inter-Event Delta_t against Event Time" )

plot( head( event_data$arrival_t, 10000 ), head( event_data$delta_t, 10000 ), col="blue", pch=19, xlab="Event Time in ns", ylab="Inter-Event Delta_t in ns", main="Inter-Event Delta_t against Event Time" )
plot( head( event_data$arrival_t, 1000 ), head( event_data$delta_t, 1000 ), col="blue", pch=19, xlab="Event Time in ns", ylab="Inter-Event Delta_t in ns", main="Inter-Event Delta_t against Event Time" )
plot( head( event_data$arrival_t, 100 ), head( event_data$delta_t, 100 ), col="blue", pch=19, xlab="Event Time in ns", ylab="Inter-Event Delta_t in ns", main="Inter-Event Delta_t against Event Time" )
plot( head( event_data$arrival_t, 10 ), head( event_data$delta_t, 10 ), col="blue", pch=19, xlab="Event Time in ns", ylab="Inter-Event Delta_t in ns", main="Inter-Event Delta_t against Event Time" )

plot( head( event_data$delta_t, 1000 ), head( event_data$duration, 1000 ), pch=19, col="blue", xlab="Inter-Event Delta_t", ylab="Duration of Event in ns", main="Event Duration against Inter-Event Delta_t" )

```

## Summary Stats and Metrics for `delta_t`.
```{r}

. <- event_data %>%
  slice( -1 ) %>%
  mutate( delta_t = as.numeric( delta_t ) ) %>%
  select( delta_t )
delta_t = .$delta_t

hist( delta_t, col="blue" )

str(delta_t)
summary(delta_t)
describe(delta_t) 
acf(delta_t)
pacf(delta_t)
```
# H0 (null): The data was produced in a random manner.
# Ha (alternative): The data was not produced in a random manner.

# w/ p-value < 2.2e-16; THere is sufficien evidence to reject the Null Hypothesis. 

d1 <- msgw_in$inter_message_delta_t_ns
d2 <- msgw_in$message_processing_time_ns

runs.test( d1 ) 
chisq.test( d1 )
bartels.rank.test( d1 )
cox.stuart.test( d1 )
difference.sign.test( d1 )
acf( d1 ) 
pacf( d1 )

runs.test( d2 ) 
chisq.test( d2 )
bartels.rank.test( d2 )
cox.stuart.test( d2 )
difference.sign.test( d2 )
acf( d2 ) 
pacf( d2 )

    # Strip Charts at different times across the data set
    arrival_times_slice_labels <- cut( numeric_arrival_times, breaks=30*60, labels=F, include.lowest=T )
    arrival_time_slices <- data.frame( numeric_arrival_times, arrival_times_slice_labels )
    par(mfrow=c( 30, 1 ), mai=c(0.1, 0.1, 0.1, 0.1) )
    loop_count <- 1:30
    for( i in loop_count ){
      stripchart( arrival_time_slices$numeric_arrival_times[ arrival_time_slices$arrival_times_slice_labels == i * 60 ], method="stack", pch=19, cex=0.5, col="blue", xlab=NULL, ylab=NULL, main=NULL, axes=F )
    }


EIplot( msgw_in$message_arrival_t_ns )
EIglm( numeric_arrival_times )
EIglm( arrival_time_slices$numeric_arrival_times[ arrival_time_slices$arrival_times_slice_labels == 1 ] ) 

```