---
title: "Data Prep"
output: html_notebook
---

```{r}
#pacman::p_load(pacman, rio, modules, data.table, bit64, dplyr, ggplot2, psych, randtests, eventInterval )
#p_unload( all )
library( tidyverse )
```

## Import `msgw_in` data from csv source.
```{r}
.msgw_in_import <- rio::import( "./Data/eqfa-msgw-in-0709.txt", format="csv", setclass="tibble" )
```

## Prepare working data

- Normalize message arrival time-stamp to start series at $t_0 = 0$.
- Calculate inter_message_delta_t in nanoseconds
- Convert latency/processing time to nanoseconds
- Calculate message exit time in nanoseconds

```{r}
t0 = first(.msgw_in_import$ts_ns)
msgw_in <- tibble( 
  message_arrival_t_ns = as.numeric( .msgw_in_import$ts_ns - t0 ),
  inter_message_delta_t_ns = message_arrival_t_ns - lag( message_arrival_t_ns, default=first(message_arrival_t_ns) ),
  message_processing_delta_t_ns = as.numeric( na_if( .msgw_in_import$"PNQM Latency (ms)", "null" ) ) * 1e6,
  message_exit_t_ns = message_arrival_t_ns + message_processing_delta_t_ns,
  message_type = .msgw_in_import$`Msg Type`
)
```

## Remove incomplete records and duplicated arrival times.

We need to gain some understanding of how the source data is generated.  These 'missing' values may or may not make sense.  For now we just remove them to make progress on the mathematics.
```{r}
msgw_in <- msgw_in %>%
  filter( !is.na( message_processing_delta_t_ns ))
```

## Reduce data down to just a couple of minutes
```{r}
msgw_in_trunc <- msgw_in[ msgw_in$message_arrival_t_ns < 1.2e11, ]
```

Successfully loaded `r nrow( msgw_in_trunc )` observations over the first two minutes of our primary data.

## Move this stuff to a Verification or Analysis Rmd
```{r}
hist( msgw_in_trunc$message_arrival_t_ns, breaks=120, col="blue", xlab = "Message Arrival Time in ns over the first two minutes of data.", main="Histogram of Message Arrival Times (bin count = 120)" )

plot( msgw_in_trunc$message_arrival_t_ns, msgw_in_trunc$inter_message_delta_t_ns, col="blue", xlab="Message Arrival Time in ns", ylab="Inter-Message Delta_t in ns", main="Inter-message delta_t against Message Arrival Time" )

stripchart( msgw_in_trunc$message_arrival_t_ns, method="stack", col="blue", xlab="Message Arrival Time in ns", main="1-D Stripchart of Message Arrival Time" )
```

# TODO
# Calculate Queue Size

# Calculate 'Leaky Integrator' or other smoothing window function/kernel

## Analysis

msgw_in_length <- length( msgw_in$message_arrival_t_ns )
msgw_in_unique_length <- length( unique( msgw_in$message_arrival_t_ns ) )
print( paste( "msgw_in$ts_nw is Unique:", msgw_in_length == msgw_in_unique_length ) )
print( paste( "Percent of msgw_in Unique arrival time: ", msgw_in_unique_length / msgw_in_length ) )

msgw_in_duplicate_arrival_time <- msgw_in[duplicated(msgw_in$message_arrival_t_ns),]
## Visual inspection of the observations with duplicate time stamps did not reveal any obvious patters.  We will leave the duplicate timestamps as they are.

print( paste( "t0 is Monotonic: ", all(msgw_in$message_arrival_t_ns == cummax(msgw_in$message_arrival_t_ns) ) ) )

print( paste( "Data covers ", max( msgw_in$message_arrival_t_ns), 'ns' ) )
print( paste( "Data covers ", max( msgw_in$message_arrival_t_ns)/1e6, 'ms' ) )
print( paste( "Data covers ", max( msgw_in$message_arrival_t_ns)/1e9, 'sec' ) )
print( paste( "Data covers ", max( msgw_in$message_arrival_t_ns)/1e9/60, 'min' ) )

# Our data covers ~30 minutes.
ggplot( msgw_in, aes( x = inter_message_delta_t_ns ) ) + geom_histogram()

hist( msgw_in$inter_message_delta_t_ns,
      col="thistle",
      main = "Histogram of Inter-message time Between Message Arrivals",
      xlab = "Nanoseconds between messages"
)
              mean( msgw_in$inter_message_delta_t_ns )
              sd( msgw_in$inter_message_delta_t_ns)

summary( msgw_in$inter_message_delta_t_ns )
describe( msgw_in$inter_message_delta_t_ns )

ggplot( msgw_in, aes( x = message_processing_time_ns )) + geom_histogram() + scale_x_log10()

tmp <- data.frame( message_arrival_t_ns = as.numeric( msgw_in$message_arrival_t_ns ) )
str( tmp )
ggplot( tmp, aes( x = message_arrival_t_ns ) ) + geom_histogram( bins = 30*60*100 ) 
hist( tmp$message_arrival_t_ns, breaks = 30*60*100 )

# H0 (null): The data was produced in a random manner.
# Ha (alternative): The data was not produced in a random manner.

# w/ p-value < 2.2e-16; THere is sufficien evidence to reject the Null Hypothesis. 

d1 <- msgw_in$inter_message_delta_t_ns
d2 <- msgw_in$message_processing_time_ns

runs.test( d1 ) 
chisq.test( d1 )
bartels.rank.test( d1 )
cox.stuart.test( d1 )
difference.sign.test( d1 )
acf( d1 ) 
pacf( d1 )

runs.test( d2 ) 
chisq.test( d2 )
bartels.rank.test( d2 )
cox.stuart.test( d2 )
difference.sign.test( d2 )
acf( d2 ) 
pacf( d2 )

stripchart( as.numeric( msgw_in$message_arrival_t_ns[ msgw_in$message_arrival_t_ns < 1 * 60 * 1e12] ) )

numeric_arrival_times <- as.numeric( msgw_in$message_arrival_t_ns )
hist( numeric_arrival_times, breaks = 30*60*100 )
stripchart( numeric_arrival_times )

arrival_times_slice_labels <- cut( numeric_arrival_times, breaks=30*60, labels=F, include.lowest=T )
arrival_time_slices <- data.frame( numeric_arrival_times, arrival_times_slice_labels )
par(mfrow=c( 30, 1 ), mai=c(0.1, 0.1, 0.1, 0.1) )
loop_count <- 1:30
for( i in loop_count ){
  stripchart( arrival_time_slices$numeric_arrival_times[ arrival_time_slices$arrival_times_slice_labels == i * 60 ], method="stack", pch=19, cex=0.5, col="blue", xlab=NULL, ylab=NULL, main=NULL, axes=F )
}

## hawkes
# https://cran.r-project.org/web/packages/hawkesbow/index.html
# https://cran.r-project.org/web/packages/hawkesbow/hawkesbow.pdf
# https://search.r-project.org/CRAN/refmans/hawkes/html/simulateHawkes.html

plot( numeric_arrival_times, msgw_in$inter_message_delta_t_ns )

EIplot( msgw_in$message_arrival_t_ns )
EIglm( numeric_arrival_times )
EIglm( arrival_time_slices$numeric_arrival_times[ arrival_time_slices$arrival_times_slice_labels == 1 ] ) 

```