---
title: "Inital Exploratory Data Analysis"
output: html_notebook
---

In this section we will attempt to gain some insight into the general nature 
of our data from mostly a descriptive statistics perspective through the use
of some visualizate and light calculations.  We aren't trying to discover the
exact mathematics that would result in good models for our observations.
Rather we are just trying to develop a modest but reasonable understanding of
our data. These results will help us determine how to focus our efforts
as we apply more rigorous (and expensive) techniques downstream to accurately
model the features and characteristics in our observations.

## Vizualizations concerning Event Arrival Time 

```{r viz-arrival_T}
hist( as.numeric( event_data$arrival_T ), breaks=120, col="blue", xlab = "Event arrival time in ns over the first two minutes of data.", main="Histogram of Event Arrival Count per Sec" )
```
It's notable that we chose to use 120 breaks in our histogram so that the bars
in the visualization all represent a 1 second time frame.  Although we should be
careful to not infer too much from this graph in terms of patterns or distributions;
we can at least see that there is certainly some variance in the rate of events
per seconds in our initial data set.

As a quick note, I think it's important that we mentioned the hesitation to
reach any particular conclusions from our results so far.  That will be a very
common feature of this section of the project.  It's crucially important that
practitioners don't let the early (and relatively trivial) analysis overly
influence later work.  It's far to easy to later find yourself spending a lot
of time and effort on an unproductive line of inquiry because of misplaced
enthusiasm in an overvalued result in this early stage.

Next we have a quick look at a few strip charts that simply plot the one
dimensional arrival time of consecutive messages. We are looking for any breaks
in homogeneity at different time scales.

```{r}
par(mfrow=c( 4, 1 ) )#, mai=c(0.1, 0.1, 0.1, 0.1) )
stripchart( as.numeric( head( event_data$arrival_T, 10000 ) ), pch=19, cex=0.5, col="blue", xlab="Event Time in ns", main="1-D Stripchart of Event Arrival Time for the first 10000 observations." )
stripchart( as.numeric( head( event_data$arrival_T, 1000 ) ), pch=19, cex=0.5, col="blue", xlab="Event Time in ns", main="1-D Stripchart of Event Arrival Time for the first 1000 observations." )
stripchart( as.numeric( head( event_data$arrival_T, 100 ) ), pch=19, cex=0.5, col="blue", xlab="Event Time in ns", main="1-D Stripchart of Event Time; first 100 observations." )
stripchart( as.numeric( head( event_data$arrival_T, 10 ) ), pch=19, cex=0.5, col="blue", xlab="Event Time in ns", main="1-D Stripchart of Event Time; first 10 observations." )
par( mfrow=c(1, 1) )
```

Given our understanding of the domain for the events that we are studying, we do
expect to build evidence to confirm that event arrival times are 'bursty'
in nature, but with a relatively homogeneous distribution in the 'noise floor' 
for the observed arrival times. 

Each of our strip charts above plot a sample size that is an order of magnitude
smaller that the one above it.  As was expected this confirms that at the extremes
the events seem more homogeneous, but there appears to be a break in symmetry to be
discovered by looking at the appropriate scales.  It's should be noted that the
data was split on sample size, not a time domain. So the number of observations
in each plot is proportionally related to those next to it.  However, the magnitude of
time represented along the abscissa is not necessary as proportionally related.

Studying the data along relative time scales could certainly be an interesting
thing to do later.  But it must be clear that in these charts we are not doing so.

In our next series of plots we are doing something very similar but extending it out
in a second dimension.  The event points are now spread out vertically
to show the inter-event time gap immediately preceding each event.  We are again
looking for breaks in homogeneity that can help us confirm, or not, our assumption
that the events are bursty in nature.  Similar to above, we can certain
see what appears to be some level of clustering in the data.  However it's far
from clear.  It's a recognized phenomena that random and independent data
can manifest grouping and clustering patterns.  We'll need more rigorous analysis
to prove that in our system there is the inter-event relationships and
self-excitation that would result in the grouping and clustering of our events.
And even more analysis to ultimately parameterize any observable excitation into a model.

```{r}
plot( head( event_data$arrival_T, 10000 ), head( event_data$delta_t, 10000 ), col="blue", pch=19, xlab="Event Time in ns", ylab="Inter-Event Delta_t in ns", main="Inter-Event Delta_t against Event Time" )
plot( head( event_data$arrival_T, 1000 ), head( event_data$delta_t, 1000 ), col="blue", pch=19, xlab="Event Time in ns", ylab="Inter-Event Delta_t in ns", main="Inter-Event Delta_t against Event Time" )
plot( head( event_data$arrival_T, 100 ), head( event_data$delta_t, 100 ), col="blue", pch=19, xlab="Event Time in ns", ylab="Inter-Event Delta_t in ns", main="Inter-Event Delta_t against Event Time" )
plot( head( event_data$arrival_T, 10 ), head( event_data$delta_t, 10 ), col="blue", pch=19, xlab="Event Time in ns", ylab="Inter-Event Delta_t in ns", main="Inter-Event Delta_t against Event Time" )
```
In any case, we can definitely see some break in homogeneity in the data.  Certainly
nothing that would imply any confidence that it's something more than chance, but
at least enough that we can leave this line of study open for closer inspection
downstream.


## Visualizations concering inter-event delta_t
```{r viz-delta_t}
. <- event_data %>%
  #slice( -1 ) %>%
  mutate( delta_t = replace_na( as.numeric( delta_t ), 0 ) ) %>%
  select( delta_t )
delta_t = .$delta_t

str(delta_t)
summary(delta_t)
describe(delta_t) 
acf(delta_t)
pacf(delta_t)

plot( event_data$arrival_T, delta_t, col="blue", pch=19, xlab="Event Time in ns", ylab="Inter-Event Delta_t in ns", main="Inter-Event Delta_t against Event Time" )

plot( head( delta_t, 1000 ), head( event_data$duration, 1000 ), pch=19, col="blue", xlab="Inter-Event Delta_t", ylab="Duration of Event in ns", main="Event Duration against Inter-Event Delta_t" )

hist( as.numeric( delta_t), col="blue", main="Histogram of Inter-event Delta_t", xlab="Delta_t in Nanoseconds")
hist( log( as.numeric( delta_t)), col="blue", main="Histogram of log( Inter-event Delta_t )", xlab="log(Delta_t) in Nanoseconds")

{
  qqnorm( as.numeric( delta_t ) )
  qqline(as.numeric(delta_t), distribution=qnorm)
}
{
  qqplot(x=qexp(ppoints(1000)), y=as.numeric(delta_t), main="Exponential Q-Q Plot",
         xlab="Theoretical Quantiles", ylab= "Delta_t Quantiles")
  qqline(as.numeric(delta_t), distribution=qexp)
}
```

## Visualizations concering the Number of Concurrent Events over Time (Queueing Process)
```{r explore_concurent_events}
in_queue <- event_data %>%
  filter( !is.na( departure_T ), departure_T != 0 ) %>%
  mutate( t = arrival_T, step = 1 ) %>%
  select( t, step )

out_queue <- event_data %>%
  filter( !is.na( departure_T ), departure_T != 0 ) %>%
  mutate( t = departure_T, step = -1 ) %>%
  select( t, step )

queue <- bind_rows( in_queue, out_queue ) %>%
  arrange( t ) %>%
  mutate( size = cumsum( step ))
  
plot( queue$t, queue$size, type="l", col="blue", pch=19, cex=0.5, main="Concurrent Event Count at Time t.", xlab='Time t in Nanoseconds', ylab="Concurrent event count" )
hist( log( queue$size ), col="blue", main="Histogram of log( Concurrent Event Count )", xlab="log( Concurrent Event Count )")
```
<!-- Scratch Notes Below
# H0 (null): The data was produced in a random manner.
# Ha (alternative): The data was not produced in a random manner.

# w/ p-value < 2.2e-16; THere is sufficien evidence to reject the Null Hypothesis. 

d1 <- msgw_in$inter_message_delta_t_ns
d2 <- msgw_in$message_processing_time_ns

runs.test( d1 ) 
chisq.test( d1 )
bartels.rank.test( d1 )
cox.stuart.test( d1 )
difference.sign.test( d1 )
acf( d1 ) 
pacf( d1 )

runs.test( d2 ) 
chisq.test( d2 )
bartels.rank.test( d2 )
cox.stuart.test( d2 )
difference.sign.test( d2 )
acf( d2 ) 
pacf( d2 )

    # Strip Charts at different times across the data set
    arrival_Times_slice_labels <- cut( numeric_arrival_Times, breaks=30*60, labels=F, include.lowest=T )
    arrival_Time_slices <- data.frame( numeric_arrival_Times, arrival_Times_slice_labels )
    par(mfrow=c( 30, 1 ), mai=c(0.1, 0.1, 0.1, 0.1) )
    loop_count <- 1:30
    for( i in loop_count ){
      stripchart( arrival_Time_slices$numeric_arrival_Times[ arrival_Time_slices$arrival_Times_slice_labels == i * 60 ], method="stack", pch=19, cex=0.5, col="blue", xlab=NULL, ylab=NULL, main=NULL, axes=F )
    }


EIplot( msgw_in$message_arrival_T_ns )
EIglm( numeric_arrival_Times )
EIglm( arrival_Time_slices$numeric_arrival_Times[ arrival_Time_slices$arrival_Times_slice_labels == 1 ] ) 
-->